### 2024-03-27 MS5130_A3_Clothing_Product_Analysis.qmd
Insights 1.1
Insights and analysis from dataset 1 and reading the other 2 dataset

---
title: "MS5130 A3 Clothing_Product_Analysis"
author: "Manav"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

## Product analysis using Dataset 1

```{r}

# Loading all the required libraries

library(ggplot2) # For visualization
library(dplyr) # For data manipulation
library(tidyr) # For tidying data
library(stringr) # For string manipulation
library(readr) # For reading CSV files
library(tidytext)
library(plotly)
library(wordcloud)
library(quanteda)
```

```{r}

# Reading the CSV File
product_catalog <- read_csv("C:/Users/manav/OneDrive/Desktop/Work/R_Applied_Analytics/Assignment 3/myntra_products_catalog.csv")

head(product_catalog)
nrow(product_catalog)

# Pre-processing data
product_catalog <- product_catalog %>%
  # Renaming columns named weirdly
  rename(PriceINR = `Price (INR)`) %>%
  # Removing rows with blank entries or NAs in any column
  ####filter(apply(., 1, function(row) all(row != "" & !is.na(row)))) %>%
  drop_na() %>%
  # Trimming leading and trailing whitespaces in all character columns
  # Adapted from stackoverflow: https://stackoverflow.com/questions/20760547/removing-whitespace-from-a-whole-data-frame-in-r
  mutate(across(where(is.character), trimws))

head(product_catalog)

```

### Analysis 1: Expensive and budget-friendly brands

```{r}

# Calculating the average price for each brand
brand_prices <- product_catalog %>%
  group_by(ProductBrand) %>%
  # Adapted from stackoverflow : https://stackoverflow.com/questions/25759891/dplyr-summarise-each-with-na-rm
  summarise(AveragePrice = mean(PriceINR, na.rm = TRUE),NumProducts = n()) %>%
  filter(NumProducts > 10) %>%
  ungroup() %>%
  arrange(desc(AveragePrice))

# Identifying the top 3 most expensive brands and cheap brands
top_expensive_brands <- head(brand_prices, 3)

top_cheap_brands <- tail(brand_prices, 3)

# Displaying the results
print("Top 3 Most Expensive Brands: ") 
print(top_expensive_brands)

print("Top 3 Budget Friendly Brands: ")
print(top_cheap_brands)

```

### Analysis 2: Gender wise product distribution

```{r}

# Getting product counts grouped by gender
gender_counts <- product_catalog %>%
  group_by(Gender) %>%
  summarise(Count = n()) %>%
  ungroup()

# Creating a bar graph for the number of products by gender
ggplot(gender_counts, aes(x = Gender, y = Count, fill = Gender, label = Count)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  theme_minimal() +
  labs(title = "Gender wise product distribution",
       x = "Gender",
       y = "Number of Products") +
  scale_fill_brewer(palette = "Set1") +
  geom_text(position = position_stack(vjust = 0.5))

```

### Analysis 3: Color wise products

```{r}

# Getting product counts for each color
color_counts <- product_catalog %>%
  group_by(PrimaryColor) %>%
  summarise(Count = n()) %>%
  arrange(desc(Count))
nrow(color_counts)

# Identifying top 5 colors
top_colors <- head(color_counts, 5)

# Putting the rest of the colors in "Others" category
others_count <- sum(tail(color_counts, n = nrow(color_counts) - 5)$Count)
others_row <- data.frame(PrimaryColor = "Others", Count = others_count)

# Combining the top colors with "Others"
final_data <- rbind(top_colors, others_row)

final_data

# setting color styles manually
# Adapted from open ai : 'Prompt - Color codes for the below colors'
named_colors <- c(
  "Black" = "#808080",  # Lighter grey instead of black
  "Blue" = "#ADD8E6",   # Light Blue
  "Red" = "#FF9999",    # Light Red
  "Green" = "#90EE90",  # Light Green
  "White" = "#F8F8FF",  # Off-White, Ghost White
  "Others" = "#D8BFD8"  # Light Purple (Thistle)
)

# Changing the 'final_data' values as percentage of whole count
final_data$Percentage <- (final_data$Count / sum(final_data$Count)) * 100

ggplot(final_data, aes(x = "", y = Count, fill = PrimaryColor, label = sprintf("%1.1f%%", Percentage))) +
  geom_bar(width = 1, stat = "identity", color = "white") +
  coord_polar(theta = "y") +
  theme_void() +
  labs(title = "Color Based Product Distribution") +
  scale_fill_manual(values = named_colors) +
  geom_text(position = position_stack(vjust = 0.5))

# blue, black and red are the most popular color categories
```

## Sentiment and Review analysis using Dataset 2 and 3

```{r}

# Reading the other 2 datasets (Detailed Product dataset and Product Review dataset)

Product_Detailed_2 <- read_csv("C:/Users/manav/OneDrive/Desktop/Work/R_Applied_Analytics/Assignment 3/Product_Detailed_2.csv")
Product_Reviews_3 <- read_csv("C:/Users/manav/OneDrive/Desktop/Work/R_Applied_Analytics/Assignment 3/Product_Reviews_3.csv")

# Manually creating the mapping based on domain knowledge so that the Categories from both the dataset can be matched to each other

Product_Detailed_2 <- Product_Detailed_2 %>%
  mutate(Category = case_when(
    Category == "Inner Wear & Sleep Wear" ~ "Intimate",
    Category == "Lingerie & Sleep Wear" ~ "Intimate",
    Category == "Western" ~ "Dresses",
    Category == "Bottom Wear" ~ "Bottoms",
    Category == "Topwear" ~ "Tops",
    Category == "Sports Wear" ~ "Jackets",
    Category == "Indian Wear" ~ "Trend",
    TRUE ~ as.character(Category)  # Default case to keep original Category values
  ))
```

###